Color-dual NLSM multi-loop integrands

### 2023-08-02 notes/2loopNLSM.tex
1-loop YZ reminder
Rearranged things, notes to self

\abstract{ For now, let's put the things we need to check/compute here in the abstract. 
\begin{itemize}
\item (1) Does our construction manifest all the cuts for the two-loop n-gon?
\item  (2) Is there a natural extension to 3-loop 
\item (3)  Is there a simple Feynman rule/ghost we can add to correct the one-loop n-gon of YZ theory
\item (4) Can we make our two-loop numerators functionally symmetric 
\item (5) What is the physics story here? Perturbative corrections to special Galilon, DBI for inflation? 
\item (6) What is the rule for cuts we don't care about? Like the weird snaking diagrams that James identified?
\item (7) How can bubble on external legs that are zero by IBP inform generalizations beyond pions?
\end{itemize}
}

@@

@@

\section{Two-loop}
\subsection{Cut construction, graph sym, Jacobi -- James}

\subsection{Pions -- James}
Let's start with a two-loop four-point example. At this order, the basis diagrams for color-dual representations are any two of the Jacobi triple, the double box, the penta-triangle or the cross-box. First we introduce a bit of shorthand notion:
\be
[LR] \equiv  l_L^2 -l_R^2 \qquad [M] \equiv  l_M^2
\ee
Now let's start with the double-box. Below we treat the color legs as $Y$ particles that will get dressed with factors of $[LR]$. Thus, we wish to manifest the following cut:
\be
N^{\text{2box}}_M\equiv \dBox{}{hred}{hred}{hred}{hred}{hred}{hred} = [12][23][34][45][56][61]
\ee
In addition, we want the following cuts to be simultaneously satisfied:
\be
N^{\text{2box}}_R\equiv \dBoxR{hred}{}{hred}{}{hred}{hred}{hred} = [61][12][27][76][4]^2
\ee
\be
N^{\text{2box}}_L\equiv \dBoxL{hred}{hred}{}{hred}{}{hred}{hred}=[45][57][73][34][1]^2
\ee
Where the middle cut takes $l_7^2\rightarrow 0$, the right cut takes $l_3^2,l_5^2\rightarrow 0$ and the left cut takes $l_2^2,l_6^2\rightarrow 0$. Applying these cuts to all the diagrams above, we obtain the following expressions:
\be
\text{Cut}_R(N^{\text{2box}}_M) = \Delta_R + N^{\text{2box}}_R \qquad \text{Cut}_L(N^{\text{2box}}_M) = \Delta_L + N^{\text{2box}}_L
\ee
However, we find that the remainder terms $\Delta_{L/R}$ are proportional to $l_7^2$, and thus vanish on the middle cut:
\be
\text{Cut}_M(\Delta_{L/R})  = 0
\ee
Similarly, applying the complement left/right cuts to the right/left remainder terms, we obtain a single kinematic function:
\be
\text{Cut}_L(\Delta_R) = \text{Cut}_R(\Delta_L) = \Delta_{LR}
\ee
Thus, the double box numerator that simultaneously manifests all of the cut contributing to NLSM are as follows:
\be
\boxed{N^{\text{NLSM}}_{\text{2box}} = N^{\text{2box}}_M - (\Delta_{L}+\Delta_{R})+\Delta_{LR}}
\ee
Applying the cut relations above, we find that this double box numerator factorizes to the desired expressions:
\be
\text{Cut}_{X}(N^{\text{NLSM}}_{\text{2box}} )= N^{\text{2box}}_{X}
\ee
The kinematic expressions obtained for the remainder functions are provided below:
\begin{align}
\Delta_L &= \frac{1}{2}[61][12]([27]+[67])[4]^2[7] 
\\
 \Delta_R &= \frac{1}{2}[34][45]([37]+[57])[1]^2[7]
 \\
\Delta_{LR}&= [1]^2[4]^2[7]^2
\end{align}
While these numerators are not functionally automorphic like the one-loop $n$-gon numerators, they do obey color-kinematics when evaluated on-shell and thus should be valid representations for double copy construction.

@@

\section{Notes}\label{sec:Notes}

\begin{itemize}
\item ZM and 2D NLSM (PCM) are the same classically but not quantum mechanically.
This observation is due to other people but an easy way to see it.
ZM and 2D NLSM have the same EOM after a field redefinition.
This means that the classical physics is the same but of course the Lagrangians aren't the same and the field redefinition can change the path integral measure.
\item Forcing pions to match ZM in 2D does funny things.
The pion coupling constant is dropped but is implicitly set by the ordered 4pt amplitude which is normalized to be $-k_1 \cdot k_3$ in the package.
I've tabulated what happens when you take a general D NLSM answer and force it to match ZM in 2D.
Let's say that $n_\text{NLSM} = z n_\text{ZM}$ in 2D where I'm matching basis/master numerators.
For 4pt tree, $z=2/3$ (I matched the half ladder).
For 4pt 1-loop, $z$ is free/pure generalized gauge (I matched the box).
For 4pt 2-loop (when you impose certain $\ell$ power counting) $z$ is $216/565=\tfrac{2^3 3^3}{5 \cdot 113}$ (I matched the double box and penta-triangle).
For 6pt tree, $z$ is 12/35 (I matched the half ladder).
So you can match kinematic structure (the combination of Mandelstams) of NLSM to ZM in 2D but you can't get the coupling constant right at loop level, that is, ZM is a different theory quantum mechanically.
Actually these numbers make it look like you can't even match the theories classically but I guess for the theories to match classically you only need the amplitudes to agree and I matched numerators.
\item At 4pt 1-loop $\prod p_L^2 -p_R^2$ agrees with the ZM answer $\prod p_L \eps p_R$ up to some number.
At 6pt 1-loop this is not true and the two answers differ by alternating signs depending on the kinematic configuration.
\item You cannot lift ZM to general D NLSM in the democratic way that Nic described because it fails at 6pt tree.
\item To clean up the NLSM 4pt 2-loop answer I forced the answer to match ZM in 2D.
The ZM answer also only has certain powers of loop momenta $\ell^5$...$\ell^8$ where $\ell$ represents any combination of $\ell_1$ and $\ell_2$.
I enforce this $\ell^5$...$\ell^8$ behavior on the NLSM ansatz too.
\end{itemize}

>>>>

\abstract{ For now, let's put the things we need to check/compute here in the abstract. 
\begin{enumerate}
\item Does our construction manifest all the cuts for the two-loop n-gon?
\item What is the physics story here? Perturbative corrections to special Galilon, DBI for inflation? 
\item How can bubble on external legs that are zero by IBP inform generalizations beyond pions?
\end{enumerate}
}

@@

Note from James:  I've explicitly checked the nice form of the 1-loop numerators at 4pt, 6pt, and 8pt.

@@

\section{Two-loop}

Notes to myself: 
\begin{itemize}
\item ZM and 2D NLSM (PCM) are the same classically but not quantum mechanically.
This observation is due to other people but an easy way to see it.
ZM and 2D NLSM have the same EOM after a field redefinition.
This means that the classical physics is the same but of course the Lagrangians aren't the same and the field redefinition can change the path integral measure.
\item Forcing pions to match ZM in 2D does funny things.
The pion coupling constant is dropped but is implicitly set by the ordered 4pt amplitude which is normalized to be $-k_1 \cdot k_3$ in the package.
I've tabulated what happens when you take a general D NLSM answer and force it to match ZM in 2D.
Let's say that $n_\text{NLSM} = z n_\text{ZM}$ in 2D where I'm matching basis/master numerators.
For 4pt tree, $z=2/3$ (I matched the half ladder).
For 4pt 1-loop, $z$ is free/pure generalized gauge (I matched the box).
For 4pt 2-loop (when you impose certain $\ell$ power counting) $z$ is $216/565=\tfrac{2^3 3^3}{5 \cdot 113}$ (I matched the double box and penta-triangle).
For 6pt tree, $z$ is 12/35 (I matched the half ladder).
So you can match kinematic structure (the combination of Mandelstams) of NLSM to ZM in 2D but you can't get the coupling constant right at loop level, that is, ZM is a different theory quantum mechanically.
Actually these numbers make it look like you can't even match the theories classically but I guess for the theories to match classically you only need the amplitudes to agree and I matched numerators.
\item At 4pt 1-loop $\prod p_L^2 -p_R^2$ agrees with the ZM answer $\prod p_L \eps p_R$ up to some number.
At 6pt 1-loop this is not true and the two answers differ by alternating signs depending on the kinematic configuration.
\item You cannot lift ZM to general D NLSM in the democratic way that Nic described because it fails at 6pt tree.
\item To clean up the NLSM 4pt 2-loop answer I forced the answer to match ZM in 2D.
The ZM answer also only has certain powers of loop momenta $\ell^5$...$\ell^8$ where $\ell$ represents any combination of $\ell_1$ and $\ell_2$.
I enforce this $\ell^5$...$\ell^8$ behavior on the NLSM ansatz too.
\end{itemize}

\subsection{Cut construction, graph sym, Jacobi -- James}

Color-dual integrands are typically generated by constructing an ansatz and then imposing various constraints.
Color-kinematics duality requires writing the integrand in terms of cubic graphs.
Any theory can be written this way by artificially inserting ratios of propagators (can blow up YM or NLSM).
Obviously the ansatz must have the correct power counting and external states.
We will take NLSM as our example theory but later we will discuss YM.
Since pions are scalars the numerators will not carry any spinors or polarization vectors.
Every vertex in NLSM scales as $k^2$ so, for example, the numerator of a cubic 2-loop diagram scales as $k^12$.
There are three conditions that are imposed on the ansatz:  (1) the integrand must have the correct unitarity cuts, (2) numerators must respect graph symmetries, and (3) Jacobi relations amongst graphs must be satisfied.
The integrand must be written in terms of cubic graphs in order to apply the Jacobi relations.
This means that there are naively X cubic 4pt 2-loop diagrams.
However, many of these diagrams are related by Jacobi identities.
Give an example.
Through repeated application of the Jacobi identities, all of the cubic diagrams can be written in terms of the double box and penta-triangle.
The basis choice is not unique.
For example, any pair of the double box, penta-triangle and crossed box would work.
So you have to have two numerator ansatze, one for the double box and one for the crossed box.
What we are describing here is global color-kinematics duality.
It is possible to give each cubic graph its own numerator and then only impose Jacobi on the cuts (see ...).
It is not always possible to achieve global CK, but when it is doable, it is typically much easier than doing Jacobi on the cuts because the latter requires far more ansatz parameters.
Anyway, now the numerator of any diagram can be written in terms of the two basis numerators (double box and crossed box).
This is part of imposing Jacobi.
There are additional Jacobi relations that impose actual constraints on the ansatze, for example ASDF.
We allow BEL graphs to have nonzero numerators since they do not contribute to physical unitarity cuts.
Jacobi relations involving BEL graphs or tadpoles are simply not solved.

Requiring numerators to respect their graph symmetries is very convenient but is not completely necessary in order to perform the double copy.
At tree level, finding manifestly color dual numerators is often achieved at the expense of manifest Bose symmetry (see YZ theory, j-theory, Alex numerators from CHY, Henrik numerators).
Nonetheless, for a pure ansatz construction it is very reasonable to impose graph symmetries.
Note that you impose that the numerator of \emph{every} graph respects its graph automorphisms.
So the graph syms of the double box impose constraints, but, for example, the crossed box imposes constraints on the ansatz numerators too because the numerator of the crossed box can be written in terms of the two basis numerators.

Finally we have to impose unitarity.
We only do physical cuts, no aesthetic cuts.
The two most important cuts for pions (or any EMU) are ...
However there are other cuts that must be imposed as well.

References for Jacobi, unitarity, etc.

After you've imposed all of the essential stuff you get X free parameters...
There is lots of generalized gauge freedom, clearly there's room to impose extra constraints.
For N=4 the extra conditions are things like no triangles, strip off a factor of stAtree (which fails at 5-loops), manifest power counting, etc.
The obvious choice here is to match to ZM
There's one known theory that has manifest color-kinematics, manifest Bose sym, and NLSM power counting and that's ZM theory.

\subsection{Pions -- James}

@@
